---
title: 'AAEC 4984/5984 -- Applied Economic Forecasting'
        
author: __Master Key__
date: "Homework #1 -- Spring 2020"
output: 
  pdf_document:
          keep_tex: no
header-includes:
 \usepackage{float}
geometry: margin=0.8in
---

```{r setup, include=FALSE}
library(fpp2)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = 'H') #Fixes the position of my graphs to where I included them in the document
```
The purpose of this assignment is to enhance your understanding of time series graphics and data pattern. It is intended to be rather straightforward and simple.

**Instructions**: In all cases, please ensure that your graphs and visuals have properly titles and axes labels, where necessary. Recall that you can use `help()` to find out about the data in each series. For your convenience, I have posted my R markdown file on our course website so that you can open and alter as you see fit. Refer to the output, whenever appropriate, when discussing the results

# Question 1: Visualizing Time Series Data

Create time plots of the following time series: `bicoal`, `chicken`, `dole`, `usdeaths`, `lynx`, `goog`, `writing`, `fancy`, `a10`, `h02`. 

- To allow multiple graphs on your page, please arrange your plots as grids. Below, I have provided the base code to achieve this (Note: you will need to install the `gridExtra` package before calling the `grid.arrange` command):


```{r Q1plots,  include=TRUE, fig.align="center",fig.show='hold',fig.height=9,out.width="\\textwidth",,tidy=TRUE, tidy.opts=list(width.cutoff=80)}
#Please feel free to edit accordingly!
g1 <- autoplot(bicoal) + ggtitle("Annual bituminous coal production") + labs(x ="" , y = "  ")
g2 <- autoplot(chicken) + ggtitle("Price of chicken in US ") + labs(x ="" , y = "  ")
g3 <- autoplot(dole) + ggtitle("Monthly total of people on unemployment benefits in Australia")+labs(x ="" , y = "  ")
g4 <- autoplot(usdeaths) + ggtitle("Monthly accidental deaths in USA") + labs(x ="" , y = "  ")
g5 <- autoplot(lynx) + ggtitle("Annual bituminous coal production")+ labs(x ="" , y = "  ")
g6 <- autoplot(goog) + ggtitle("Daily closing stock prices of Google") + labs(x ="" , y = "  ")
g7 <- autoplot(writing) + ggtitle("Sales for printing and writing paper") + labs(x ="" , y = "  ")
g8 <- autoplot(fancy) + ggtitle("Monthly sales for a souvenir shop") + labs(x ="" , y = "  ")
g9 <- autoplot(a10) + ggtitle("Monthly anti-diabetic drug sales in Australia") + labs(x ="" , y = "  ")
g10 <- autoplot(h02) + ggtitle("Monthly corticosteroid drug sales") + labs(x ="" , y = "  ")

gridExtra::grid.arrange(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10, nrow=5,ncol=2,newpage=TRUE)
```


# Question 2: Assessing Seasonality

i. Use the `ggseasonplot()`, `ggsubseriesplot()`, and `ggAcf`\footnote{It might be useful to set the max lag in the ACF to 36 so that you can see a fair bit of the patterns in the correlogram.} functions to explore possible seasonality in the following time series: `writing`, `fancy`, `a10`, `h02`.

```{r Q2plots, include=TRUE, fig.align="center",fig.height=6, out.width="\\textwidth", fig.cap="Seasonal plot of selected series",tidy=TRUE,tidy.opts=list(width.cutoff=70)}
gg1 <- ggseasonplot(writing) + theme(legend.position="bottom", legend.text=element_text(size=6))
gg2 <- ggseasonplot(fancy) + theme(legend.position="bottom", legend.text=element_text(size=6))
gg3 <- ggseasonplot(a10) + theme(legend.position="bottom", legend.text=element_text(size=6))
gg4 <- ggseasonplot(h02) + theme(legend.position="bottom", legend.text=element_text(size=6))
gridExtra::grid.arrange(gg1,gg2,gg3,gg4, nrow=2,ncol=2)
```
```{r Q2plotsb, include=TRUE, fig.align="center",fig.height=6, out.width="\\textwidth", fig.cap="Subseries plot of selected series",tidy=TRUE,tidy.opts=list(width.cutoff=70)}
gg1 <- ggsubseriesplot(writing)
gg2 <- ggsubseriesplot(fancy) 
gg3 <- ggsubseriesplot(a10) 
gg4 <- ggsubseriesplot(h02) 
gridExtra::grid.arrange(gg1,gg2,gg3,gg4, nrow=2,ncol=2)
```
```{r Q2plotsc, include=TRUE, fig.align="center",fig.height=6, out.width="\\textwidth", fig.cap="ACF plot of selected series",tidy=TRUE,tidy.opts=list(width.cutoff=70)}
gg1 <- ggAcf(writing,lag.max=36) 
gg2 <- ggAcf(fancy,lag.max = 36) 
gg3 <- ggAcf(a10,lag.max = 36) 
gg4 <- ggAcf(h02,lag.max = 36) 
gridExtra::grid.arrange(gg1,gg2,gg3,gg4, nrow=2,ncol=2)
```


ii. What can you say about the seasonal patterns?

**Answer: From the plots, we notice that there is clear seasonality across all three chart types. In Figure \ref{fig:Q2plots}, we notice that the month of August is persistently lowest for the writing series. This is confirmed by Figure \ref{fig:Q2plotsb} where the mean for the month of August is lower than all the other months. The `ho2` series has a similar pattern for the month of February but trends upwards for all the other months. The ACFs in Figure \ref{fig:Q2plotsc} confirms that there is indeed seasonality at the monthly frequency because there is a spike at the 12th (and multiples of 12) lag in all the data series. Our `a10` series displays a strong trend as well.**

iii. Can you identify any unusual years?

**Answer: In December, 1992 the monthly sales for a souvenir shop increased dramatically compared to the same month of the last year. We could see this a bit easier if we included the `polar=TRUE` command in the `ggseasonplot`. It is OK if you didn't notice this!**

# Question 3: White Noise

`dj` contains 292 consecutive trading days of the Dow Jones Index. 

i. Plot this series and its ACF. Comment on any pattern noticed in both. Does this series look like white noise?

```{r Q3plot, include=TRUE, fig.show = "hold", out.width="\\textwidth", fig.cap= "Time Plot \\& ACF of the Dow-Jones Index"}
dj1 <- autoplot(dj,col="blue4") + ggtitle("Daily Dow-Jones index") + labs(x="",y="")
dj2 <- ggAcf(dj, col="red") + ggtitle("ACF: Daily Dow-Jones index")

gridExtra::grid.arrange(dj1,dj2)
```

**Answer: There is a general upward trend in the Dow-Jones Index (see Figure \ref{fig:Q3plot}). This is confirmed by the slowly decaying ACF in Figure \ref{fig:Q3plot}. Clearly, this series is not WN.**

ii. Now, use `ddj <- diff(dj)` to compute the daily changes in the index. 
```{r}
ddj <- diff(dj)
```


iii. Plot `ddj` and its ACF. Do the changes in the Dow Jones Index look like white noise?

```{r Q3plotb, include=TRUE, fig.show = "hold", out.width="\\textwidth", fig.cap= "Time Plot \\& ACF of the Differenced Dow-Jones Index"}
ddj1 <- autoplot(ddj,col="blue4") + ggtitle("Differenced Daily Dow-Jones index") + labs(x="",y="")
ddj2 <- ggAcf(ddj, col="red") + ggtitle("ACF: Differenced Daily Dow-Jones index")

gridExtra::grid.arrange(ddj1,ddj2)
```

**Answer: Figure \ref{fig:Q3plotb} indicates that the series can be WN. More than 95% of the autocorrelations lie within the CI.**
